Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-05-19 doc/mc-paper.tex
Incremental changes

%\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

@@

  For Boolean formula $\phi$ and ring weight function $w$, the \emph{ring evaluation problem} computes
  \begin{eqnarray}
    \rep(\phi, w) & = & \sum_{\alpha \in \modelset(\phi)} \;\; \prod_{\lit \in \alpha} w(\ell) \label{eqn:rep}
  \end{eqnarray}
  In this equation, sum \scalebox{0.8}{$\sum$} is computed using addition operation $\radd$, and product \scalebox{0.8}{$\prod$} is computed using multiplication operation $\rmul$.
\label{def:weight}
\end{definition}

@@

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
\label{prop:ring:eval}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any weight function $w$:
\begin{eqnarray}
%% V1
\textstyle
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Pand_{1 \leq i \leq k} \phi_i,\; w \right) &=& \prod_{1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Por_{1 \leq i \leq k} \phi_i,\; w\right) &=& \sum_{1 \leq i \leq k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}
As is described in \cref{section:formally-verified-toolchain}, we have proved these three equations using \lean{}.

@@

As notation,
let $\psi$ be a subset of the clauses in $\inputformula$.
For partial assignment
$\passign$, the expression  $\simplify{\psi}{\passign}$ denotes the set of clauses $\gamma$
obtained from $\psi$ by: i) eliminating any
clause containing a literal $\lit$ such that $\passign(\lit) = 1$,
ii) for the remaining clauses eliminating those literals $\lit$ for
which $\passign(\lit) = 0$, and iii) eliminating any duplicate clauses.
In doing these simplifications, we also track the \emph{provenance}
of each simplified clause $C$, i.e., which of the (possibly multiple) input clauses simplified to become $C$.
More formally, for $C \in \simplify{\psi}{\passign}$, we let $\prov_{\passign}(C, \psi)$ denote
those clauses $C' \in \psi$, such that
$C' \subseteq C \cup \bigcup_{\lit \in \passign} \obar{\lit}$.
We then extend the definition of $\prov$ to any simplified formula
$\gamma$ as $\prov_{\passign}(\gamma, \psi) = \bigcup_{C \in \gamma} \prov_{\passign}(C, \psi)$.

>>>>

\Copyright{Randal E. Bryant, Wojciech Nawrocki, Jeremy Avigad, and Marijn J. H. Heule} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

@@

  For Boolean formula $\phi$ and ring weight function $w$, the \emph{ring evaluation problem} computes
  \begin{equation}
    \begin{array}{rcl}
    \rep(\phi, w) & = & \sum_{\alpha \in \modelset(\phi)} \;\; \prod_{\lit \in \alpha} w(\ell) \label{eqn:rep}
    \end{array}
  \end{equation}
  In this equation, sum \scalebox{0.8}{$\sum$} is computed using addition operation $\radd$, and product \scalebox{0.8}{$\prod$} is computed using multiplication operation $\rmul$.
\label{def:weight}
\end{definition}

@@

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
\label{prop:ring:eval}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any weight function $w$:
%%\begin{eqnarray}
\begin{displaymath}
\begin{array}{rcl}
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \\ %\label{eqn:ring:negation} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Pand_{1 \leq i \leq k} \phi_i,\; w \right) &=& \prod_{1 \leq i \leq k} \rep(\phi_i, w) \\ %\label{eqn:ring:product} \\
\textstyle
\scalebox{1.2}{\rep}\left(\Por_{1 \leq i \leq k} \phi_i,\; w\right) &=& \sum_{1 \leq i \leq k} \rep(\phi_i, w) \\ %\label{eqn:ring:sum}
\end{array}
\end{displaymath}
\end{proposition}
As is described in \cref{section:formally-verified-toolchain}, we have proved these three equations using \lean{}.

@@

As notation,
let $\psi$ be a subset of the clauses in $\inputformula$.
For partial assignment
$\passign$, the expression  $\simplify{\psi}{\passign}$ denotes the set of clauses $\gamma$
obtained from $\psi$ by: i) eliminating any
clause containing a literal $\lit$ such that $\passign(\lit) = 1$,
ii) for the remaining clauses eliminating those literals $\lit$ for
which $\passign(\lit) = 0$, and iii) eliminating any duplicate or tautological clauses.
In doing these simplifications, we also track the \emph{provenance}
of each simplified clause $C$, i.e., which of the (possibly multiple) input clauses simplified to become $C$.
More formally, for $C \in \simplify{\psi}{\passign}$, we let $\prov_{\passign}(C, \psi)$ denote
those clauses $C' \in \psi$, such that
$C' \subseteq C \cup \bigcup_{\lit \in \passign} \obar{\lit}$.
We then extend the definition of $\prov$ to any simplified formula
$\gamma$ as $\prov_{\passign}(\gamma, \psi) = \bigcup_{C \in \gamma} \prov_{\passign}(C, \psi)$.
